\documentclass[12pt,a4paper,oneside]{scrreprt}

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage[english]{babel}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{bm}
\usepackage{booktabs}
\usepackage{csquotes}
\usepackage{graphicx}
\usepackage{microtype}
\usepackage{siunitx}
\usepackage{subcaption}
\usepackage{xcolor}
\usepackage{hyperref}

\hypersetup{
  colorlinks=true,
  linkcolor=blue!50!black,
  citecolor=blue!50!black,
  urlcolor=blue!50!black
}

\graphicspath{{figures/}{../figures/}}

\newcommand{\SU}{\mathrm{SU}}
\newcommand{\U}{\mathrm{U}}
\newcommand{\dd}{\mathrm{d}}
\newcommand{\order}{\mathcal{O}}
\newcommand{\eps}{\varepsilon}
\newcommand{\latspacing}{a}
\newcommand{\wilsonloop}{W(R,T)}
\newcommand{\creutz}{\chi(R,T)}
\newcommand{\rzero}{r_0}
\newcommand{\stringtension}{\sigma}

\title{Investigation of the Continuum Limit in the Presence of Small Breaking of Gauge Invariance in Lattice Gauge Theories}
\author{Piet Leyendecker}
\date{\today}

\begin{document}

\maketitle

\begin{abstract}
This thesis investigates whether and how the continuum limit of a lattice gauge
theory is affected by small explicit violations of gauge invariance. The study
combines theoretical scaling arguments with numerical measurements of Wilson
loops, Creutz ratios, the static potential, the Sommer scale, and related
dimensionless observables. The central goal is to determine whether
gauge-breaking effects vanish, persist, or modify the extrapolation toward the
continuum limit as the lattice spacing is reduced.
\end{abstract}

\tableofcontents

\chapter*{Notation and Conventions}
\addcontentsline{toc}{chapter}{Notation and Conventions}
\begin{description}
  \item[$\latspacing$] Lattice spacing.
  \item[$\beta$] Bare lattice coupling parameter.
  \item[$\eps_1,\eps_2$] Bare parameters controlling the explicit breaking of gauge invariance.
  \item[$\wilsonloop$] Wilson loop with spatial extent $R$ and temporal extent $T$.
  \item[$\creutz$] Creutz ratio used to estimate the force and string tension.
  \item[$\rzero$] Sommer scale defined through $r^2F(r)=1.65$.
  \item[$\stringtension$] String tension.
\end{description}

\chapter{Introduction}
\label{chap:introduction}

\section{Context and Problem}
Gauge invariance is one of the defining structural principles of lattice gauge
theory. In Wilson's formulation it is preserved exactly at nonzero lattice
spacing, which strongly constrains the allowed lattice artifacts and supports a
controlled approach to the continuum limit. In practical and generalized
settings, however, gauge invariance may be only approximate. Small explicit
violations can arise from regulator artifacts, effective descriptions,
algorithmic imperfections, or imperfect implementations of gauge constraints.

The central problem of this thesis is therefore not whether exact gauge
invariance is useful, but how robust the continuum limit is when the lattice
theory is perturbed by a controlled, small gauge-breaking deformation. This
requires a comparison between the standard gauge-invariant theory and deformed
ensembles along a sequence of lattice spacings.

\section{Research Gap}
The theoretical expectation depends on the scaling character of the
gauge-breaking deformation. If the deformation acts like an irrelevant lattice
artifact, its effect should vanish as the lattice spacing is reduced. If it
contains a relevant component, it may change the continuum theory unless it is
tuned away. The numerical study must therefore distinguish a harmless change of
cutoff effects from a genuine shift of continuum observables.

\section{Aim and Research Questions}
The aim of the thesis is to test this distinction numerically for pure
$\SU(2)$ lattice gauge theory with a small explicit gauge-breaking deformation.
The guiding research question is:
\begin{quote}
  Does a small explicit breaking of gauge invariance disappear in the continuum limit, or does it alter the continuum extrapolation of physical observables?
\end{quote}

\section{Strategy and Main Observables}
The numerical strategy is to compare gauge-invariant and gauge-broken ensembles
through observables that have a well-defined continuum interpretation. The main
observables are temporal Wilson loops, Creutz ratios, the static potential
$V(R)$, the Sommer scale $\rzero$, and the string tension $\stringtension$ or
dimensionless combinations derived from it. These quantities provide both scale
setting information and direct probes of the continuum extrapolation.

\section{Structure of the Thesis}
Chapter~\ref{chap:theoretical-background} summarizes the theoretical
background needed for the analysis. Chapter~\ref{chap:small-breaking} derives
the concrete research gap and formulates the gauge-breaking framework and
working hypotheses. Chapter~\ref{chap:methodology} describes the simulation
setup, observable extraction, continuum extrapolation, and uncertainty
estimation. Chapter~\ref{chap:results} presents the numerical results.
Chapter~\ref{chap:discussion} interprets the findings in relation to the
research question and the theoretical expectations. The final chapter
summarizes the conclusions and identifies directions for further research.

\chapter{Scientific Background and Theoretical Framework}
\label{chap:theoretical-background}

The aim of this chapter is to collect the theoretical ingredients needed for the
analysis of the continuum limit in the presence of small explicit violations of
gauge invariance. The discussion starts from continuum Yang--Mills theory and
then introduces the lattice formulation, Wilson loops, the static potential,
Creutz ratios, and the Sommer scale. These are the central concepts used later
to define and compare physical observables at different lattice spacings.

\section{Gauge Theories in the Continuum}
\label{sec:continuum-gauge-theory}

\subsection{Gauge Fields and Local Gauge Symmetry}
\label{subsec:gauge-fields-local-symmetry}

A non-Abelian gauge theory is built from fields that transform locally under a
Lie group. In this thesis the relevant case is the group $\SU(2)$. A continuum
gauge field can be written as
\begin{equation}
  A_\mu(x) = A_\mu^a(x) T^a ,
\end{equation}
where $T^a$ are the generators of the Lie algebra of $\SU(2)$. A common choice is
$T^a = \sigma^a/2$, with $\sigma^a$ the Pauli matrices. The gauge field defines
the covariant derivative
\begin{equation}
  D_\mu = \partial_\mu + i A_\mu ,
\end{equation}
and the associated field-strength tensor
\begin{equation}
  F_{\mu\nu}
  =
  \partial_\mu A_\nu
  -
  \partial_\nu A_\mu
  +
  i [A_\mu,A_\nu] .
\end{equation}
The commutator term distinguishes a non-Abelian gauge theory from an Abelian
one. It implies that the gauge field itself carries charge under the gauge
group and therefore self-interacts.

A local gauge transformation is specified by a group element
\begin{equation}
  \Omega(x) \in \SU(2)
\end{equation}
at every point in space-time. Under such a transformation, matter fields in the
fundamental representation transform as
\begin{equation}
  \psi(x) \mapsto \Omega(x)\psi(x),
\end{equation}
while the gauge field transforms inhomogeneously,
\begin{equation}
  A_\mu(x)
  \mapsto
  \Omega(x) A_\mu(x) \Omega^\dagger(x)
  -
  i \bigl(\partial_\mu \Omega(x)\bigr)\Omega^\dagger(x).
\end{equation}
The inhomogeneous derivative term is necessary because the transformation is
local. The field strength transforms covariantly,
\begin{equation}
  F_{\mu\nu}(x)
  \mapsto
  \Omega(x) F_{\mu\nu}(x) \Omega^\dagger(x),
\end{equation}
so traces of products of field strengths at the same point are gauge invariant.

In Euclidean space-time, the pure Yang--Mills action can be written as
\begin{equation}
  S_{\mathrm{YM}}
  =
  \frac{1}{2g^2}
  \int \dd^4 x \,
  \mathrm{tr}\,
  F_{\mu\nu}(x) F_{\mu\nu}(x),
  \label{eq:continuum-yang-mills-action}
\end{equation}
with summation over repeated Lorentz indices. This action is invariant under
local gauge transformations. The coupling $g$ controls the strength of the
interaction. For an asymptotically free gauge theory, the effective coupling
becomes small at short distances, while the long-distance theory is strongly
coupled and requires non-perturbative methods.

\subsection{Gauge-Invariant Observables}
\label{subsec:gauge-invariant-observables}

Physical observables in a gauge theory must be gauge invariant. Gauge-dependent
quantities can be useful after a gauge fixing has been imposed, but they do not
directly correspond to physical observables unless the gauge dependence is
properly controlled. For pure gauge theory, important gauge-invariant
observables include traces of closed parallel transporters, local action-density
operators, correlation functions of gauge-invariant operators, and quantities
derived from Wilson loops.

A central continuum object is the path-ordered parallel transporter along a
curve $C$,
\begin{equation}
  U(C)
  =
  \mathcal{P}
  \exp\left(
    i \int_C \dd x^\mu A_\mu(x)
  \right).
\end{equation}
For a closed curve, the trace
\begin{equation}
  W(C)
  =
  \frac{1}{N_c}
  \mathrm{Re}\,\mathrm{tr}\, U(C)
\end{equation}
is gauge invariant. On the lattice this object becomes the Wilson loop and is
one of the most direct probes of confinement and of the static quark--antiquark
potential.

\subsection{Physical Consequences of Gauge Invariance}
\label{subsec:consequences-gauge-invariance}

Gauge invariance is not only a formal redundancy of the description. It
constrains the allowed terms in the action, the form of correlation functions,
and the possible continuum limits. In the lattice formulation, preserving gauge
invariance exactly at nonzero lattice spacing is usually advantageous because it
prevents gauge-noninvariant counterterms from being generated by radiative
corrections. Wilson's lattice gauge action realizes this idea by putting compact
group-valued variables on links and by constructing the action from traces of
closed loops \cite{Wilson1974,GattringerLang2010,MontvayMuenster1994}.

The purpose of this thesis is to study what happens when this exact symmetry is
slightly violated by an explicit deformation. From a renormalization-group point
of view, the key question is whether the deformation behaves like an irrelevant
lattice artifact, so that its effects vanish as $\latspacing \to 0$, or whether
it defines a relevant perturbation that changes the continuum limit. This
question is not purely formal: small violations of gauge constraints can appear
as regulator artifacts, algorithmic imperfections, or implementation errors in
effective or quantum-simulation formulations of lattice gauge theories
\cite{Foerster1980,Bonati2021Breaking}.

\section{Lattice Gauge Theory}
\label{sec:lattice-gauge-theory}

\subsection{Discretization of Space-Time}
\label{subsec:discretization}

In lattice gauge theory, Euclidean space-time is replaced by a hypercubic
lattice
\begin{equation}
  \Lambda_{\latspacing}
  =
  \left\{
    x = \latspacing n
    \,\middle|\,
    n \in \mathbb{Z}^4
  \right\},
\end{equation}
where $\latspacing$ is the lattice spacing. The lattice spacing acts as an
ultraviolet cutoff. Momenta larger than order $1/\latspacing$ are not resolved.
The continuum limit is obtained by sending
\begin{equation}
  \latspacing \to 0
\end{equation}
while keeping physical quantities fixed.

In a numerical simulation the lattice has finite extent
\begin{equation}
  L_\mu = N_\mu \latspacing ,
\end{equation}
where $N_\mu$ is the number of lattice points in direction $\mu$. Therefore two
limits are conceptually distinct: the continuum limit $\latspacing \to 0$ and
the infinite-volume limit $L_\mu \to \infty$. In practice, one tries to choose
volumes large enough that finite-volume effects are smaller than the statistical
and systematic errors relevant for the analysis. In the simulation campaign of
this thesis, spatial volumes are organized to keep the physical spatial extent
approximately fixed where possible, while the temporal extent is chosen as part
of the ensemble definition and analysis strategy.
[Check against the actual simulation setup: in the simulations in this thesis I
use $L_\mu=\mathrm{const}$ for the spatial directions and
$N_\mu=\mathrm{const}$ for the time direction.]

\subsection{Link Variables and Plaquettes}
\label{subsec:link-variables-plaquettes}

The fundamental degrees of freedom of Wilson lattice gauge theory are link
variables
\begin{equation}
  U_\mu(x) \in \SU(2),
\end{equation}
which live on the oriented link from $x$ to $x+\latspacing\hat\mu$. They are the
lattice analogues of parallel transporters. For a smooth continuum field one may
formally write
\begin{equation}
  U_\mu(x)
  \approx
  \exp\left[
    i \latspacing A_\mu\left(x+\frac{\latspacing}{2}\hat\mu\right)
  \right].
\end{equation}
A local lattice gauge transformation assigns a group element $\Omega(x)$ to each
site and acts on the link variables as
\begin{equation}
  U_\mu(x)
  \mapsto
  \Omega(x) U_\mu(x) \Omega^\dagger(x+\latspacing\hat\mu).
  \label{eq:lattice-gauge-transformation}
\end{equation}

The elementary closed loop on the lattice is the plaquette,
\begin{equation}
  U_{\mu\nu}(x)
  =
  U_\mu(x)
  U_\nu(x+\latspacing\hat\mu)
  U_\mu^\dagger(x+\latspacing\hat\nu)
  U_\nu^\dagger(x).
  \label{eq:plaquette-definition}
\end{equation}
Under a gauge transformation, the plaquette transforms covariantly at its base
point,
\begin{equation}
  U_{\mu\nu}(x)
  \mapsto
  \Omega(x) U_{\mu\nu}(x) \Omega^\dagger(x).
\end{equation}
Therefore $\mathrm{Re}\,\mathrm{tr}\,U_{\mu\nu}(x)$ is gauge invariant.

For smooth fields, the plaquette has the continuum expansion
\begin{equation}
  U_{\mu\nu}(x)
  =
  \exp\left[
    i \latspacing^2 F_{\mu\nu}(x)
    +
    \order(\latspacing^3)
  \right].
\end{equation}
Thus the plaquette is the lattice object from which the continuum field
strength is recovered. Although the later analysis focuses on Wilson loops,
static potentials, and Creutz ratios, this relation explains why the Wilson
plaquette action reproduces Yang--Mills theory in the continuum limit.

\subsection{Wilson Gauge Action}
\label{subsec:wilson-gauge-action}

The standard Wilson gauge action for $\SU(N_c)$ is
\begin{equation}
  S_W[U]
  =
  \beta
  \sum_x
  \sum_{\mu<\nu}
  \left(
    1
    -
    \frac{1}{N_c}
    \mathrm{Re}\,\mathrm{tr}\,
    U_{\mu\nu}(x)
  \right),
  \label{eq:standard-wilson-action}
\end{equation}
with
\begin{equation}
  \beta = \frac{2N_c}{g_0^2}.
\end{equation}
For $\SU(2)$ this becomes
\begin{equation}
  \beta = \frac{4}{g_0^2},
\end{equation}
and
\begin{equation}
  S_W[U]
  =
  \beta
  \sum_x
  \sum_{\mu<\nu}
  \left(
    1
    -
    \frac{1}{2}
    \mathrm{Re}\,\mathrm{tr}\,
    U_{\mu\nu}(x)
  \right).
  \label{eq:wilson-action-su2}
\end{equation}
The action is built only from traces of closed loops and is therefore exactly
gauge invariant at nonzero lattice spacing.

The continuum limit of the Wilson action is obtained by inserting the plaquette
expansion. Up to an irrelevant additive constant and lattice artifacts,
\begin{equation}
  S_W[U]
  \longrightarrow
  \frac{1}{2g_0^2}
  \int \dd^4x\,
  \mathrm{tr}\,
  F_{\mu\nu}(x)F_{\mu\nu}(x).
\end{equation}
This reproduces the Euclidean Yang--Mills action. The corrections at finite
lattice spacing are cutoff effects. For the pure Wilson gauge action, the
leading gauge-invariant lattice artifacts in many on-shell quantities are of
order $\latspacing^2$, although the actual scaling observed in numerical data
can be affected by statistics, finite-volume effects, operator definitions, and
the accessible range of lattice spacings.

\subsection{Path Integral and Expectation Values}
\label{subsec:lattice-path-integral}

The lattice path integral is finite-dimensional at finite lattice spacing and
finite volume. The partition function is
\begin{equation}
  Z
  =
  \int
  \prod_{x,\mu}
  \dd U_\mu(x)
  \,
  e^{-S[U]},
  \label{eq:lattice-partition-function}
\end{equation}
where $\dd U_\mu(x)$ is the Haar measure on $\SU(2)$. The expectation value of
an observable $\mathcal{O}[U]$ is
\begin{equation}
  \langle \mathcal{O} \rangle
  =
  \frac{1}{Z}
  \int
  \prod_{x,\mu}
  \dd U_\mu(x)
  \,
  \mathcal{O}[U]
  e^{-S[U]}.
  \label{eq:lattice-expectation-value}
\end{equation}
Gauge-invariant expectation values can be computed without gauge fixing. This
is one of the practical advantages of the compact Wilson formulation.

In Monte Carlo simulations the path integral is estimated by generating an
ensemble of gauge-field configurations distributed according to $e^{-S[U]}$.
For an ensemble of $N_{\mathrm{cfg}}$ configurations $U_i$, the estimator is
\begin{equation}
  \langle \mathcal{O} \rangle
  \approx
  \frac{1}{N_{\mathrm{cfg}}}
  \sum_{i=1}^{N_{\mathrm{cfg}}}
  \mathcal{O}[U_i].
\end{equation}
Because successive configurations in a Markov chain are correlated,
autocorrelations must be considered when estimating uncertainties.

\subsection{Continuum Limit and Renormalization}
\label{subsec:continuum-limit-renormalization}

The continuum limit is not obtained by simply decreasing $\latspacing$ at fixed
bare coupling. Instead, the bare parameters have to be tuned so that physical
observables remain fixed in physical units. For pure $\SU(2)$ gauge theory with
the Wilson action, decreasing the lattice spacing corresponds to increasing
$\beta$, or equivalently decreasing the bare coupling $g_0$.

A useful way to formulate the continuum limit is in terms of a physical length
scale, such as the Sommer scale $\rzero$. The dimensionless ratio
\begin{equation}
  \frac{\rzero}{\latspacing}
\end{equation}
is measured on the lattice. Increasing $\rzero/\latspacing$ means that the same
physical length is resolved by more lattice points, i.e. the lattice spacing is
smaller in physical units. Therefore continuum extrapolations can be expressed
as fits in
\begin{equation}
  \left(\frac{\latspacing}{\rzero}\right)^2
\end{equation}
or another measured dimensionless lattice-spacing proxy.

Continuum extrapolations compare observables at fixed physical
conditions, not merely at fixed lattice indices. For observables derived from
Wilson loops this means that a loop size $R$ on a coarse ensemble and the same
integer $R$ on a fine ensemble do not represent the same physical distance.
Therefore one should compare at fixed ratios such as $r/\rzero$ or other fixed
dimensionless distances. This point is emphasized in recent work on controlled
continuum extrapolations of Wilson-loop observables, where Creutz ratios are
used as finite continuum quantities and fitted at fixed physical distance
\cite{Risch2024SmearingContinuum}.

The renormalization-group interpretation is central for the present thesis.
Different lattice actions can lead to the same continuum theory if they differ
only by irrelevant operators. In that case they lie in the same universality
class. By contrast, a relevant deformation changes the long-distance theory
unless its coefficient is tuned to zero. A gauge-breaking term must therefore be
classified by its effect on long-distance, gauge-invariant observables as
$\latspacing \to 0$. The numerical analysis in later chapters tests this
question directly.

\section{Static Potential and Confinement}
\label{sec:static-potential-confinement}

\subsection{Wilson Loops}
\label{subsec:wilson-loops}

A Wilson loop is the trace of the ordered product of link variables around a
closed contour. For a rectangular loop in the $\mu\nu$-plane with side lengths
$L_\mu$ and $L_\nu$ in lattice units, the code measures the normalized real
trace
\begin{equation}
  W_{\mu\nu}(L_\mu,L_\nu)
  =
  \frac{1}{N_c V}
  \sum_x
  \operatorname{Re}\operatorname{tr}
  \left[
    U_\mu(x;L_\mu)\,
    U_\nu(x+L_\mu\hat\mu;L_\nu)\,
    U_\mu^\dagger(x+L_\nu\hat\nu;L_\mu)\,
    U_\nu^\dagger(x;L_\nu)
  \right],
  \label{eq:rectangular-wilson-loop-code}
\end{equation}
where $V$ is the lattice volume and
\begin{equation}
  U_\mu(x;L_\mu)
  =
  U_\mu(x)
  U_\mu(x+\hat\mu)
  \cdots
  U_\mu(x+(L_\mu-1)\hat\mu)
\end{equation}
denotes a straight Wilson line of length $L_\mu$ in direction $\mu$. The factor
$1/N_c$ normalizes the trace, while the factor $1/V$ averages over all possible
base points of the loop.

For temporal Wilson loops, the last lattice direction is used as the Euclidean
time direction,
\begin{equation}
  t = N_d - 1.
\end{equation}
The code computes rectangular loops in each spatial-time plane and averages over
the spatial directions:
\begin{equation}
  W(R,T)
  =
  \frac{1}{N_d-1}
  \sum_{\mu=0}^{N_d-2}
  W_{\mu,t}(R,T).
  \label{eq:temporal-wilson-loop-spatial-average}
\end{equation}
For the $\SU(2)$ simulations considered in this thesis, $N_c=2$, so the stored
temporal Wilson loop is
\begin{equation}
  W(R,T)
  =
  \frac{1}{2V(N_d-1)}
  \sum_{\mu=0}^{N_d-2}
  \sum_x
  \operatorname{Re}\operatorname{tr}
  W_{\mu,t}(x;R,T).
  \label{eq:temporal-wilson-loop-su2-code}
\end{equation}
Thus the values written to the output file are already normalized by the color
factor, averaged over the full lattice volume, and averaged over spatial
directions.

In the implementation, the rectangular loop is constructed from two paths that
connect opposite corners of the rectangle. The first path is
\begin{equation}
  l_\mu
  =
  U_\mu(x;L_\mu)\,
  U_\nu(x+L_\mu\hat\mu;L_\nu),
\end{equation}
and the second path is
\begin{equation}
  l_\nu
  =
  U_\nu(x;L_\nu)\,
  U_\mu(x+L_\nu\hat\nu;L_\mu).
\end{equation}
The measured loop contribution is then
\begin{equation}
  \operatorname{Re}\operatorname{tr}
  \left(
    l_\mu l_\nu^\dagger
  \right).
\end{equation}
This is equivalent to the trace of the ordered product around the rectangular
closed contour.

The temporal Wilson-loop output file has the structure
\begin{verbatim}
# step, L, T, W_temp
\end{verbatim}
where \texttt{L} is the spatial extent $R$, \texttt{T} is the temporal extent,
and \texttt{W\_temp} is the averaged, normalized value defined in
Eq.~\eqref{eq:temporal-wilson-loop-spatial-average}.

\subsection{Temporal Smearing and Flow-Time Convention}
\label{subsec:temporal-smearing-flow-time}

The signal-to-noise problem for temporal Wilson loops becomes more severe as
the loop area increases. The simulations and analysis therefore support two
separate smoothing procedures. The first is a multihit estimator for the links
entering the temporal Wilson-loop measurement. The second is measurement on a
gauge field that has first been evolved by Wilson, or gradient, flow. These two
procedures are conceptually distinct: multihit improves the estimator for a
given ensemble, while gradient flow changes the gauge field on which the
observable is evaluated.

For a temporal loop with spatial extent $L$ and temporal extent $T$, the
measured object can be written as
\begin{equation}
  W(L,T)
  =
  \frac{1}{N_c}
  \operatorname{Re}\operatorname{tr}
  \left[
    S_i(x,L)\,
    T_t(x+L\hat i,T)\,
    S_i^\dagger(x+T\hat t,L)\,
    T_t^\dagger(x,T)
  \right],
  \label{eq:temporal-wilson-loop-transporters}
\end{equation}
followed by the average over all lattice sites $x$ and all spatial directions
$i$. Here $S_i(x,L)$ is the straight spatial transporter of length $L$ in
direction $i$, and $T_t(x,T)$ is the straight temporal transporter of length
$T$. In the implementation the temporal direction is the last lattice
direction, $t=N_d-1$.

The raw measurement constructs the required spatial transporters up to the
largest requested spatial length. The temporal transporter is then built
iteratively,
\begin{equation}
  T_t(x,T+1)
  =
  T_t(x,T)\,
  U_t(x+T\hat t).
  \label{eq:temporal-transporter-recursion}
\end{equation}
For each requested pair $(L,T)$, the closed loop in
Eq.~\eqref{eq:temporal-wilson-loop-transporters} is accumulated over all
origins and spatial orientations. The value written to \texttt{W\_temp.out} is
normalized by
\begin{equation}
  (N_d-1)\,V\,N_c,
\end{equation}
where $V$ is the lattice volume. This is the same normalization used in
Eq.~\eqref{eq:temporal-wilson-loop-spatial-average}.

When multihit temporal smearing is enabled, links entering the loop estimator
are replaced by local averages generated with the same local action used for
the simulation update. For Metropolis simulations, the local average is
estimated from repeated Metropolis proposals for the selected link at fixed
surrounding staple. For heatbath simulations, the averaged link is generated
from local heatbath updates, optionally followed by the configured number of
overrelaxation steps. Setting
\begin{equation}
  \texttt{wilson\_loop\_multihit}=1
\end{equation}
disables this link averaging and gives the raw Wilson-loop estimator.

Gradient flow is handled separately. For each measured configuration $U$, a
copy $V(t)$ is evolved by Wilson flow using a third-order Runge--Kutta
integrator. The flow force is constructed from the local staple, and after
each flow step the links are projected back onto the gauge group. Wilson loops
at positive flow time are measured on the flowed links $V_\mu(x,t)$, not on the
original links $U_\mu(x)$.

The flow time stored in the output is the dimensionless lattice flow time
\begin{equation}
  t_{\mathrm{lat}}
  =
  \frac{t}{\latspacing^2}.
  \label{eq:dimensionless-flow-time}
\end{equation}
Accordingly, the smoothing radius in lattice units is
\begin{equation}
  \frac{\sqrt{8t}}{\latspacing}
  =
  \sqrt{8t_{\mathrm{lat}}}.
  \label{eq:flow-smoothing-radius}
\end{equation}
The gradient-flow output files therefore label flow time as
\texttt{t\_over\_a2}. Temporal Wilson loops measured during flow are written to
\texttt{gradient\_flow\_wtemp.dat} with the column structure
\begin{verbatim}
# conf_id, t_over_a2, L, T, W_temp
\end{verbatim}
where \texttt{conf\_id} is the saved-configuration identifier and
\texttt{W\_temp} is the normalized temporal Wilson loop at the specified flow
time.

At flow time $t=0$, the configured multihit estimator is still used. For
positive flow time, the code sets the Wilson-loop multihit count to one, so
the smoothing of the measured observable comes only from the gradient flow.
This convention avoids combining a flowed observable with an additional
stochastic link-averaging procedure.

\subsection{Effective Potential Extraction}
\label{subsec:effective-potential-extraction}

Temporal Wilson loops are used to extract the static quark--antiquark potential.
For large Euclidean time extent $T$, the Wilson loop has the spectral form
\begin{equation}
  W(R,T)
  =
  \sum_n c_n(R) e^{-a E_n(R) T},
  \label{eq:wilson-loop-spectral-decomposition}
\end{equation}
where $E_0(R)=V(R)$ is the static ground-state potential and the higher
$E_n(R)$ describe excited states with the same static sources. At sufficiently
large $T$, the ground state dominates:
\begin{equation}
  W(R,T)
  \propto
  e^{-a V(R) T}.
\end{equation}

A lattice effective potential is therefore defined by the ratio
\begin{equation}
  a V_{\mathrm{eff}}(R,T)
  =
  -
  \ln
  \left[
    \frac{W(R,T+1)}{W(R,T)}
  \right].
  \label{eq:effective-potential-ratio}
\end{equation}
If excited-state contributions are negligible, this approaches the desired
static potential in lattice units:
\begin{equation}
  a V_{\mathrm{eff}}(R,T)
  \xrightarrow[T\to\infty]{}
  a V(R).
\end{equation}

In the numerical analysis, $W(R,T)$ denotes the already normalized and spatially
averaged observable stored as \texttt{W\_temp}. No additional color, volume, or
spatial-direction normalization is applied during the analysis. The main
remaining analysis choices are the thermalization cut, the treatment of
autocorrelations, the temporal range used for the extraction, and the estimation
of statistical uncertainties.

\subsection{Creutz Ratios}
\label{subsec:creutz-ratios}

Creutz ratios are combinations of neighboring Wilson loops designed to cancel
some perimeter and normalization contributions. In the continuum they can be
viewed as the mixed derivative
\begin{equation}
  \chi(r,t)
  =
  -
  \frac{\partial^2}{\partial r\,\partial t}
  \ln W(r,t),
  \label{eq:continuum-creutz-ratio}
\end{equation}
where $W(r,t)$ is a rectangular Wilson loop of physical spatial extent $r$ and
Euclidean time extent $t$. A central finite-difference discretization gives a
lattice Creutz ratio located at half-integer separations,
\begin{equation}
  \chi\left(R+\frac{1}{2},T+\frac{1}{2}\right)
  =
  \ln
  \left[
    \frac{
      W(R+1,T) W(R,T+1)
    }{
      W(R,T) W(R+1,T+1)
    }
  \right],
  \label{eq:creutz-ratio-central-difference}
\end{equation}
or, equivalently after a shift of indices,
\begin{equation}
  \chi(R,T)
  =
  -
  \ln
  \left[
    \frac{
      W(R,T) W(R-1,T-1)
    }{
      W(R,T-1) W(R-1,T)
    }
  \right].
  \label{eq:creutz-ratio}
\end{equation}
The normalization of the Wilson loop, including the factor $1/N_c$, cancels in
this ratio. This is useful for the present $\SU(2)$ analysis because the same
formula can be applied directly to the normalized Wilson-loop output described
above.

The central-difference form approximates the continuum mixed derivative with
cutoff effects of order $\latspacing^2$ when the underlying Wilson loop is
smooth. This makes Creutz ratios particularly useful for continuum studies:
they are finite observables, their leading normalization and perimeter
contributions cancel, and they are closely related to the static force. In the
large-time limit,
\begin{equation}
  \frac{\chi(R,T)}{\latspacing^2}
  \xrightarrow[T\to\infty]{}
  F(r)
\end{equation}
up to the usual distinction between the discrete lattice midpoint and the
physical distance $r$. Diagonal ratios $\chi(R,R)$ are also useful summary
observables because they probe the same spatial and temporal scale and are
straightforward to compare across ensembles \cite{Creutz1980,Risch2024SmearingContinuum}.

If large Wilson loops obey an area law,
\begin{equation}
  W(R,T)
  \sim
  \exp[-\stringtension \latspacing^2 R T
  - \text{perimeter terms}],
\end{equation}
then the Creutz ratio approaches
\begin{equation}
  \chi(R,T)
  \xrightarrow[R,T\to\infty]{}
  \stringtension \latspacing^2 .
\end{equation}
Thus Creutz ratios provide a direct lattice estimator for the string tension in
lattice units.

At finite $R$ and $T$, Creutz ratios contain corrections from excited states,
Coulombic contributions, finite-volume effects, and lattice artifacts. They are
therefore best interpreted either as approximate string-tension estimators at
large loop sizes or as diagnostic observables for scaling comparisons between
ensembles.

For continuum extrapolations the important distinction is between fixed lattice
separation and fixed physical distance. Holding $R$ fixed while taking
$\latspacing\to0$ drives the physical distance $r=R\latspacing$ to zero and
therefore probes the short-distance limit rather than the same physical force.
The continuum comparison should instead use a dimensionless distance such as
\begin{equation}
  \hat r = \frac{r}{\rzero}
\end{equation}
and interpolate the measured Creutz ratios to common $\hat r$ values before
fitting in $\latspacing/\rzero$. This is also where gauge-breaking effects are
expected to be most visible: short-distance Wilson-loop observables are
sensitive to ultraviolet deformations, while large loops suffer from rapidly
increasing statistical noise.

\subsection{Sommer Scale and String Tension}
\label{subsec:sommer-scale-string-tension}

The static potential can also be used to set the scale. A common
phenomenological form is the Cornell-type potential
\begin{equation}
  V(r)
  =
  V_0
  -
  \frac{\alpha}{r}
  +
  \stringtension r,
  \label{eq:cornell-potential}
\end{equation}
where $V_0$ is an additive self-energy, $\alpha$ parametrizes the short-distance
Coulombic part, and $\stringtension$ is the string tension. The corresponding
force is
\begin{equation}
  F(r)
  =
  \frac{\dd V}{\dd r}.
\end{equation}
The additive constant $V_0$ drops out of the force.

The Sommer scale $\rzero$ is defined implicitly by
\begin{equation}
  \rzero^2 F(\rzero) = 1.65 .
  \label{eq:sommer-scale-definition}
\end{equation}
This definition uses the force at intermediate distances, where the potential
can often be determined more accurately than at very large distances. In lattice
units, the force is estimated from finite differences of the potential. If
$\hat V(R)=\latspacing V(R\latspacing)$, then
\begin{equation}
  \hat F\left(R+\frac{1}{2}\right)
  =
  \hat V(R+1) - \hat V(R)
\end{equation}
is a dimensionless estimate of $\latspacing^2 F(r)$ at an intermediate distance.
The Sommer condition can then be written as
\begin{equation}
  \left(\frac{r}{\latspacing}\right)^2
  \hat F\left(\frac{r}{\latspacing}\right)
  =
  1.65 .
  \label{eq:sommer-scale-lattice-units}
\end{equation}
Solving this equation gives $\rzero/\latspacing$.

The ratio $\rzero/\latspacing$ is one of the main scale-setting quantities in
this thesis. It allows data at different $\beta$ values to be compared in
approximately physical units and provides the continuum variable
\begin{equation}
  \frac{\latspacing}{\rzero}
  =
  \left(\frac{\rzero}{\latspacing}\right)^{-1}.
\end{equation}
The continuum limit corresponds to
\begin{equation}
  \frac{\latspacing}{\rzero}
  \to 0.
\end{equation}

The string tension may also be expressed in dimensionless form as
\begin{equation}
  \stringtension \rzero^2 .
\end{equation}
This quantity is useful because it is independent of the arbitrary choice of
lattice units. Agreement of such dimensionless quantities along a sequence of
lattice spacings is a central test of scaling toward a common continuum limit.

\chapter{Research Gap and Gauge-Breaking Framework}
\label{chap:small-breaking}

\section{Research Gap Derived from the Background}
The background chapter establishes two facts that are central for this thesis.
First, the Wilson plaquette action preserves local gauge invariance exactly at
nonzero lattice spacing. Second, the continuum limit is not tested directly on
the bare link variables, but through dimensionless observables such as Wilson
loops, Creutz ratios, the static potential, the Sommer scale, and combinations
involving the string tension. The research gap is the relation between these
two facts when the lattice action is deliberately deformed by terms that are
not locally gauge invariant.

The standard universality argument says that many changes of the lattice action
do not change the continuum theory. If the added terms correspond only to
irrelevant cutoff effects, their contribution to physical observables should
disappear as $\latspacing \to 0$. Gauge invariance, however, is a particularly
restrictive symmetry. Once it is broken at finite lattice spacing, the usual
symmetry argument that forbids gauge-noninvariant counterterms no longer
applies. A small breaking term can therefore be harmless only if its effect is
dynamically suppressed along the continuum trajectory, or if the coefficient is
tuned in such a way that the gauge-invariant fixed point is recovered.

This leaves an empirical question for the present simulations. The deformation
is known explicitly at the level of the update action, and its strength can be
varied. What is not known a priori is whether the measured gauge-invariant
observables approach the same continuum values as in the undeformed theory, or
whether the deformation shifts the continuum extrapolation. The role of the
numerical analysis is therefore not merely to observe a dependence on the
breaking parameters at fixed lattice spacing. Such a dependence is expected.
The central question is whether this dependence remains in the continuum
intercept after the data are expressed in physical units.

\section{Definition of the Gauge-Breaking Deformation}
The simulations start from the usual Wilson plaquette action and add explicit
single-link terms. Up to additive constants that do not affect expectation
values, the action used in the Metropolis path is
\begin{equation}
  S[U]
  =
  -\frac{\beta}{N_c}
  \sum_p
  \operatorname{Re}\operatorname{tr} U_p
  -
  \frac{\eps_1}{2}
  \sum_{x,\mu}
  \operatorname{Re}\operatorname{tr} U_\mu(x)
  -
  \eps_2
  \sum_{x,\mu}
  \left[
    \operatorname{Re}\operatorname{tr} U_\mu(x)
  \right]^2 .
  \label{eq:gauge-breaking-action}
\end{equation}
Here $U_p$ denotes the plaquette matrix, $N_c=2$ for the ensembles studied in
this work, and $\eps_1$ and $\eps_2$ are bare parameters controlling the
strength of the explicit breaking. The first term is the Wilson plaquette
action written without the irrelevant constant. The second and third terms are
not built from closed loops. They act directly on individual links and are
therefore sensitive to the representative chosen on each gauge orbit.

For a proposed replacement of a single link $U_{\mathrm{old}}$ by
$U_{\mathrm{new}}$, the Metropolis acceptance step uses the local action
difference
\begin{align}
  \Delta S
  &=
  -\frac{\beta}{N_c}
  \left[
    \operatorname{Re}\operatorname{tr}
    \left(U_{\mathrm{new}} S\right)
    -
    \operatorname{Re}\operatorname{tr}
    \left(U_{\mathrm{old}} S\right)
  \right]
  \notag\\
  &\quad
  -
  \frac{\eps_1}{2}
  \left[
    \operatorname{Re}\operatorname{tr} U_{\mathrm{new}}
    -
    \operatorname{Re}\operatorname{tr} U_{\mathrm{old}}
  \right]
  \notag\\
  &\quad
  -
  \eps_2
  \left[
    \left(\operatorname{Re}\operatorname{tr} U_{\mathrm{new}}\right)^2
    -
    \left(\operatorname{Re}\operatorname{tr} U_{\mathrm{old}}\right)^2
  \right],
  \label{eq:gauge-breaking-metropolis-delta-s}
\end{align}
where $S$ is the usual staple sum coupled to the updated link. This expression
is the local form of Eq.~\eqref{eq:gauge-breaking-action}. Positive $\eps_1$
favours links with large positive
$\operatorname{Re}\operatorname{tr} U_\mu(x)$ and therefore biases links toward
the identity element. The quadratic term controlled by $\eps_2$ favours links
with large
$\left|\operatorname{Re}\operatorname{tr} U_\mu(x)\right|$ in the Metropolis
implementation.

The heatbath and overrelaxation code uses a modified local matrix
\begin{equation}
  M
  =
  \beta S
  +
  \frac{N_c\eps_1}{2}\,\mathbb{1},
  \label{eq:heatbath-modified-local-matrix}
\end{equation}
so the heatbath kernel includes the linear $\eps_1$ deformation. Although the
heatbath parameter structure contains an $\eps_2$ entry, the current heatbath
kernel does not include the quadratic $\eps_2$ contribution. Consequently,
ensembles generated with the Metropolis update realize both terms in
Eq.~\eqref{eq:gauge-breaking-action}, whereas heatbath or overrelaxation
updates realize only the linear breaking term unless stated otherwise. This
algorithmic distinction is important when comparing ensembles produced by
different update paths.

The symmetry breaking follows directly from the lattice gauge transformation
law in Eq.~\eqref{eq:lattice-gauge-transformation}. The Wilson plaquette term is
invariant because each plaquette transforms by conjugation at its base point.
A single link transforms as
\begin{equation}
  U_\mu(x)
  \mapsto
  \Omega(x) U_\mu(x) \Omega^\dagger(x+\hat\mu),
\end{equation}
and therefore
\begin{equation}
  \operatorname{Re}\operatorname{tr} U_\mu(x)
  \mapsto
  \operatorname{Re}\operatorname{tr}
  \left[
    \Omega(x) U_\mu(x) \Omega^\dagger(x+\hat\mu)
  \right].
  \label{eq:single-link-trace-gauge-transform}
\end{equation}
For independent transformations at the two endpoints of the link, this is not
equal to the original trace in general. Thus nonzero $\eps_1$ or $\eps_2$
explicitly breaks the local gauge symmetry. The exact local redundancy of the
Wilson theory is replaced by a theory whose Boltzmann weight distinguishes
between different representatives of the same Wilson gauge orbit.

One exact symmetry remains. If the same group element is used at every lattice
site,
\begin{equation}
  \Omega(x)=\Omega ,
\end{equation}
then each link transforms by conjugation,
\begin{equation}
  U_\mu(x) \mapsto \Omega U_\mu(x)\Omega^\dagger ,
\end{equation}
and the trace is unchanged. The deformed action therefore preserves the global
diagonal $\SU(N_c)$ subgroup while breaking the independent site-dependent
gauge transformations.

\section{Symmetry and Scaling Expectations}
\subsection{Relevant, Marginal, and Irrelevant Breaking Terms}
The continuum interpretation of Eq.~\eqref{eq:gauge-breaking-action} depends
on how the breaking terms scale near the gauge-invariant continuum limit. In
renormalization-group language there are three qualitatively different
possibilities. If the deformation is irrelevant, it changes the finite-spacing
approach to the continuum but not the limiting theory. If it is marginal, its
effect can remain visible over a broad range of lattice spacings and may
require a more detailed analysis of logarithmic or weak scaling violations. If
it contains a relevant component, then a fixed nonzero breaking strength changes
the continuum theory unless the corresponding bare coefficient is tuned to
zero.

The single-link trace terms are local and ultra-violet in character. They act
on individual links rather than on extended physical observables. This makes it
plausible that small values of $\eps_1$ and $\eps_2$ may first appear as cutoff
distortions in short-distance quantities such as small Wilson loops and short
Creutz ratios. Plausibility, however, is not a proof of irrelevance. Since the
terms are not protected by local gauge symmetry, the numerical analysis must
allow both outcomes: a shared continuum intercept with modified cutoff effects,
or a shifted continuum intercept.

\subsection{Possible Restoration of Gauge Invariance}
Gauge invariance can be effectively restored in the continuum limit if the
breaking contribution to physical observables vanishes as
$\latspacing\to0$. In that case the deformed ensembles and the Wilson
ensembles lie on different lattice trajectories but approach the same
continuum fixed point. For a dimensionless observable $\hat O$, this situation
has the schematic form
\begin{equation}
  \hat O(\hat a,\eps_1,\eps_2)
  =
  O_0
  +
  c_a \hat a^2
  +
  c_1 \eps_1 \hat a^{p_1}
  +
  c_2 \eps_2 \hat a^{p_2}
  +
  \cdots ,
  \qquad
  p_1,p_2>0,
  \label{eq:restored-gauge-symmetry-scaling}
\end{equation}
where $\hat a=\latspacing/\rzero$. The coefficients and powers are observable
dependent, but the important point is the common intercept $O_0$.

If restoration does not occur at fixed breaking strength, the leading
dependence can instead enter the continuum intercept,
\begin{equation}
  \hat O(\hat a,\eps_1,\eps_2)
  =
  O_0
  +
  d_1\eps_1
  +
  d_2\eps_2
  +
  c_a \hat a^2
  +
  \cdots .
  \label{eq:nonrestored-gauge-symmetry-scaling}
\end{equation}
In this case the deformed theory approaches a different continuum limit, or no
controlled common continuum limit is visible in the accessible range. The
distinction between Eqs.~\eqref{eq:restored-gauge-symmetry-scaling} and
\eqref{eq:nonrestored-gauge-symmetry-scaling} is the main target of the
continuum fits in Chapter~\ref{chap:methodology}.

\subsection{Expected Dependence on Breaking Strength and Lattice Spacing}
At fixed $\beta$ and fixed volume, increasing $\eps_1$ should increase the
average link trace because the action gives lower weight to links far from the
identity. This also affects Wilson loops indirectly, since products of links
become biased toward larger traces. For sufficiently small $\eps_1$, the
leading response of a generic observable is expected to be approximately
linear unless a symmetry forbids the linear term. The $\eps_2$ term is
quadratic in the real link trace. In the Metropolis implementation it is
therefore even under
$\operatorname{Re}\operatorname{tr}U_\mu(x)\mapsto
-\operatorname{Re}\operatorname{tr}U_\mu(x)$ at the level of a single link, but
it is still gauge breaking and can shift gauge-invariant observables through
the generated ensemble distribution.

The lattice-spacing dependence is the decisive part of the test. If the
breaking terms only modify cutoff effects, then differences between broken and
unbroken ensembles should shrink when observables are compared at common
physical distance and plotted against $\hat a^2$. If the deformation changes
the continuum theory, the broken data may remain separated from the
gauge-invariant baseline as $\hat a\to0$. The methodology therefore uses two
classes of fits: shared-intercept fits, which encode restoration of gauge
invariance, and separate-intercept fits, which test for a persistent breaking
effect.

The analysis also uses the rescaled parameter
\begin{equation}
  \bar\eps
  =
  \frac{\eps_1}{\beta}
  \left(\frac{\rzero}{\latspacing}\right)^2 ,
  \label{eq:rescaled-breaking-parameter-framework}
\end{equation}
as a diagnostic for comparing ensembles at different lattice spacings. This
quantity is not a substitute for the bare action parameters in
Eq.~\eqref{eq:gauge-breaking-action}; rather, it provides a dimensionless way
to organize the observed breaking strength relative to the measured scale.

\section{Working Hypotheses}
The numerical analysis tests the following hypotheses.
\begin{enumerate}
  \item For sufficiently small $\eps_1$ and $\eps_2$, gauge-invariant observables can approach the same continuum values as in the Wilson theory, even though local gauge symmetry is broken at finite lattice spacing.
  \item At fixed lattice spacing, the deformation produces measurable shifts in short-distance observables, especially the average link trace, small Wilson loops, and small Creutz ratios.
  \item If the breaking behaves as an irrelevant cutoff effect, the shifts between broken and unbroken ensembles decrease with a positive power of $\latspacing/\rzero$ after comparison at fixed physical distance.
  \item If the deformation contains a relevant component, fixed nonzero $\eps_1$ or $\eps_2$ produces continuum extrapolations whose intercepts are incompatible with the gauge-invariant baseline.
  \item There may be a practical small-breaking regime. Beyond it, the response to $\eps_1$ or $\eps_2$ can become nonlinear, continuum fits can become unstable, and the assumptions behind a perturbative breaking analysis no longer apply.
\end{enumerate}

\chapter{Methodological Approach}
\label{chap:methodology}

\section{Simulation Setup and Parameters}

The gauge-field ensembles were generated for pure $\SU(2)$ lattice gauge theory
on four-dimensional hypercubic lattices. Each ensemble is specified by an
\texttt{input.yaml} file, which contains the lattice geometry
$N_0\times N_1\times N_2\times N_3$, the bare coupling $\beta$, the
gauge-breaking parameters $\eps_1$ and $\eps_2$, the Metropolis update
parameters, and the measurement setup. The simulations use the Wilson gauge
action supplemented by the gauge-breaking deformation introduced in
Chapter~\ref{chap:small-breaking}. Gauge configurations are generated with a
local Metropolis algorithm. One sweep consists of updating all link variables of
the lattice, with \texttt{nHits} Metropolis proposals per link. The proposal
size is controlled by the parameter \texttt{delta}. The main bare parameters
varied in the simulation campaign are the coupling $\beta$, which controls the
lattice spacing, and the breaking strength $\eps_1$. The parameter $\eps_2$ is
kept at zero unless stated otherwise. Measurements are performed after fixed
sweep intervals and include the plaquette, the average link trace, and temporal
Wilson loops. The temporal Wilson loops are measured for selected spatial and
temporal extents $(R,T)$ and are written to \texttt{W\_temp.out}; these data form
the basis for the extraction of the static potential, Creutz ratios, and scale
setting quantities.

\section{Ensemble Organization and Fixed-Physics Logic}
Explain how ensembles are grouped by common physical and algorithmic
parameters. State which parameters are varied in the continuum study and which
are held fixed. The organizing principle is that comparisons across $\beta$
should be made at fixed physical conditions as far as the available data allow,
rather than at fixed lattice indices alone.

\section{Thermalization, Autocorrelations, and Bootstrap}
\subsection{Thermalization Cuts}
Describe how the initial non-equilibrated part of each Markov chain is
identified and removed before combining runs.

\subsection{Integrated Autocorrelation Time}
Estimate or monitor autocorrelation times for representative observables in
order to choose measurement spacing and bootstrap block lengths.

\subsection{Bootstrap Blocking}

The measurements after thermalization are still not statistically independent,
because consecutive configurations are generated by a Markov chain. A naive
bootstrap over individual measurements would therefore underestimate the
statistical error. To account for this, the ordered measurement history is
divided into contiguous blocks of fixed length \(b\) in simulation time. The
bootstrap samples are then generated by drawing these blocks with replacement
and averaging over the resampled history. Unless stated otherwise, the
statistical errors quoted in this work use \(N_{\mathrm{boot}}=1000\) bootstrap
replicas.

The block length must be large compared with the autocorrelation time of the
observables used in the analysis. If \(b\) is too small, correlations remain
between the resampled data and the error estimate is biased downward. If \(b\)
is too large, the number of available blocks becomes small and the bootstrap
estimate becomes noisy. In practice, several block sizes are compared and the
chosen value is taken from the region where the estimated errors are stable
under further increases of \(b\).

The same bootstrap samples are propagated through the nonlinear analysis steps.
Wilson loops are resampled first, and derived quantities such as effective
potentials, the static potential \(V(R)\), Creutz ratios, and the Sommer scale
are recomputed for each bootstrap replica. This preserves correlations between
quantities derived from the same ensemble and gives the statistical
uncertainties used in the later continuum analysis.

\section{Analysis Pipeline and Data Reduction}
The final analysis starts from the individual simulation runs and groups
together runs with the same base parameters, in particular the same
$\beta$, lattice size, $\eps_1$, and $\eps_2$. Runs in such a group differ only
by statistical details such as the random seed or the total number of sweeps and
are treated as independent Markov chains for the same ensemble.

For each input run the early Monte Carlo history is inspected before combining
the data. In practice, Wilson loops and simple observables are plotted as a
function of the simulation step. A thermalization cutoff is chosen from these
plots. Measurements before this cutoff are discarded, and the remaining
measurements from all runs in the group are combined into one data set.

The combined post-thermalization data are then used to estimate the Wilson
loops $W(R,T)$ and their statistical uncertainties. Errors are obtained by a
blocked bootstrap, with a block size chosen large enough to reduce the effect of
autocorrelations. From the Wilson loops, effective masses are formed and fit
windows in $T$ are selected to extract the static potential $V(R)$.

The finalized potential points are fitted with a Cornell-type ansatz to obtain
the Sommer scale $\rzero/\latspacing$. This scale is then used to form the
dimensionless lattice spacing
\begin{equation}
  \hat a = \frac{\latspacing}{\rzero}
\end{equation}
and the rescaled breaking parameter
\begin{equation}
  \bar\eps =
  \frac{\eps_1}{\beta}
  \left(\frac{\rzero}{\latspacing}\right)^2 .
\end{equation}
Creutz ratios and available gradient-flow observables are evaluated from the
same combined ensemble. The resulting finalized ensemble summaries are the
inputs for the later comparison across different $\beta$ values and breaking
strengths.

\section{Extraction of Physical Observables}
\subsection{Wilson-Loop Analysis}
Describe the effective-mass plateaus, the choice of temporal fit windows, and
the systematic uncertainty associated with varying those windows.

\subsection{Static Potential}
Describe the potential extraction, Cornell-type fits, force determination, and
Sommer-scale extraction.

\subsection{Creutz-Ratio Analysis}
The finalized analysis computes Creutz ratios directly from the measured
temporal Wilson loops, not from the fitted static potential. The input data are
the post-thermalization \texttt{W\_temp} measurements of the individual
simulation runs belonging to one ensemble. For each run, the analysis reads
\texttt{W\_temp.out}. If gradient-flow Wilson loops were produced, the
additional Wilson-loop file named in \texttt{input.yaml} can also be loaded and
stored under its nonzero flow-time labels, but the default finalized
\(R=T\) Creutz-ratio plot uses the unflowed \(t_{\mathrm{flow}}=0\) Wilson loops
unless the analysis explicitly requests another flow time. Each row contains the
simulation step, the spatial loop size \(L\), the temporal loop size \(T\), and
the normalized temporal Wilson-loop value \texttt{W\_temp}. After applying the
chosen thermalization cut for each run,
the remaining measurements with the same \((L,T)\) are concatenated across all
runs in the ensemble. This combined Wilson-loop history is the only data source
used for the standard Creutz ratios in the finalized output.

For every available adjacent set of loop sizes, the code forms
\begin{equation}
  \chi\left(R+\frac{1}{2},T+\frac{1}{2}\right)
  =
  \ln
  \left[
    \frac{
      W(R,T+1) W(R+1,T)
    }{
      W(R,T) W(R+1,T+1)
    }
  \right].
\end{equation}
Thus a displayed diagonal point at \(R=T=1.5\), for example, is calculated from
the four measured Wilson loops \(W(1,1)\), \(W(1,2)\), \(W(2,1)\), and
\(W(2,2)\). More generally, the diagonal plot
\texttt{creutz\_ratios\_R\_eq\_T.html} keeps only those Creutz-ratio midpoints
for which the spatial and temporal midpoint are equal. The horizontal coordinate
is therefore the midpoint value \(R+1/2=T+1/2\), not an original integer loop
extent.

The central values are computed by first averaging each of the four required
Wilson-loop histories and then inserting these four means into the logarithmic
ratio. Statistical errors are obtained by the same blocked bootstrap used for
the Wilson-loop analysis. In each bootstrap replica, contiguous blocks of saved
configurations are resampled, the four Wilson-loop means are recomputed from
the resampled histories, and the Creutz ratio is recalculated from those
bootstrap means. The standard deviation of the resulting bootstrap distribution
is stored as the Creutz-ratio error. Bootstrap samples for each accepted ratio
are also written to the finalized analysis directory.

In the finalized output, the complete set of ratios is saved in
\texttt{creutz/summary.json} under the keys \texttt{chi}, \texttt{chi\_err},
and \texttt{chi\_bootstrap\_paths}. The keys of \texttt{chi} are midpoint pairs
such as \texttt{1.5,1.5}. The summary script used for scans then reads this
file and extracts the positive-within-error diagonal entries as columns named
\texttt{chi\_R=T\_1.5}, \texttt{chi\_R=T\_2.5}, and so on. These summary
columns are therefore derived from
the finalized Creutz-ratio JSON file, which in turn is derived from the combined
post-thermalization \texttt{W\_temp} Wilson-loop measurements of the source
runs.

\subsection{Scale Setting}
Summarize the determination of the Sommer scale in lattice units, possible
alternative scale estimates, and consistency checks.

\section{Continuum Extrapolation and Scaling Tests}
\subsection{Choice of Continuum Variable}
The primary continuum variable is chosen to be the measured dimensionless
lattice spacing
\begin{equation}
  \hat a
  =
  \frac{\latspacing}{\rzero}.
\end{equation}
For observables that are expected to have leading even-power cutoff effects
with the Wilson gauge action, the default horizontal axis is therefore
$\hat a^2=(\latspacing/\rzero)^2$. Using the measured scale rather than the bare
coupling makes the extrapolation closer to a fixed-physics comparison and
reduces the risk of interpreting changes in scale setting as physical changes
in the observable itself.

For Creutz-ratio observables the quantity to extrapolate should be
dimensionless. Convenient choices are
\begin{equation}
  \hat\chi(\hat r)
  =
  \rzero^2
  \frac{\chi(r,t)}{\latspacing^2},
  \qquad
  \hat r = \frac{r}{\rzero},
\end{equation}
or, for large loops where the string-tension interpretation is appropriate,
$\stringtension\rzero^2$. The measured lattice ratio $\chi(R,T)$ is
dimensionless and approaches $\stringtension\latspacing^2$ in the area-law
regime; division by $\latspacing^2$ and multiplication by $\rzero^2$ converts it
to a quantity with a finite physical continuum limit. The same logic applies to
the force extracted from the static potential.

\subsection{Fixed-Physics Trajectories}
Ensembles at different $\beta$ values are compared along trajectories where the
physical volume and the dimensionless analysis point are kept as fixed as the
available data allow. In practice this means comparing observables at common
$\hat r=r/\rzero$ and using only ensembles whose volumes are large enough that
finite-volume corrections are not expected to dominate the Wilson-loop signal.
When exact matching is not possible, the mismatch is included as a systematic
effect or avoided by restricting the fit range.

The gauge-breaking parameters are treated as additional bare deformation
parameters. If they are held fixed while $\latspacing$ changes, the continuum
fit tests whether their effect vanishes dynamically or survives in the
continuum intercept. If a theoretical scaling prescription for $\eps$ is used,
for example $\eps\propto \latspacing^q$, the fit must instead include this
scaling explicitly. The central diagnostic is whether the broken and unbroken
data approach a common continuum value.

\subsection{Fit Ansatz}
The fit strategy compares a gauge-invariant baseline with combined
lattice-spacing and breaking-dependence models.

\noindent\textbf{Gauge-invariant baseline.}
For the gauge-invariant ensembles, a minimal continuum ansatz for a
dimensionless observable $\hat O$ is
\begin{equation}
  \hat O(\hat a)
  =
  O_0
  +
  c_a \hat a^2
  +
  c_{a4} \hat a^4
  +
  \cdots .
  \label{eq:gauge-invariant-continuum-fit}
\end{equation}
The $\hat a^4$ term is included only when the data support it or when it is used
as a fit-systematic variation. The default fit should be stable under removing
the coarsest lattice spacing and under modest changes of the physical-distance
range used for the observable.

\medskip
\noindent\textbf{Combined lattice-spacing and breaking dependence.}
A useful shared-intercept test is
\begin{equation}
  \hat O(\hat a,\eps)
  =
  O_0
  +
  c_a \hat a^2
  +
  c_\eps \eps \hat a^p
  +
  c_{a\eps} \eps \hat a^{p+2}
  +
  \cdots ,
  \label{eq:shared-intercept-breaking-fit}
\end{equation}
with $p>0$. This ansatz encodes the hypothesis that gauge breaking is an
irrelevant cutoff effect: for any fixed small $\eps$ the breaking terms vanish
as $\hat a\to0$, and all data share the same continuum intercept $O_0$.

The competing separate-intercept test is
\begin{equation}
  \hat O(\hat a,\eps)
  =
  O_0
  +
  c_\eps \eps
  +
  c_a \hat a^2
  +
  c_{a\eps} \eps \hat a^2
  +
  \cdots .
  \label{eq:separate-intercept-breaking-fit}
\end{equation}
Here a statistically significant $c_\eps$ indicates that the breaking
deformation changes the continuum value of the observable rather than merely
changing cutoff effects. If the deformation can enter symmetrically in the sign
of $\eps$, a quadratic term $c_{\eps2}\eps^2$ should be included as a systematic
variant.

This structure follows the same logic used in controlled continuum
extrapolations with smearing or flow: one separates changes in the path toward
the continuum from changes in the continuum observable itself
\cite{Risch2024SmearingContinuum}. In that setting, fixed smearing in lattice
units can share the same continuum intercept if it remains in the controlled
expansion regime, while a fixed physical smoothing radius defines a different
continuum observable. In the present gauge-breaking problem there is no
guarantee of a shared intercept, so Eqs.~\eqref{eq:shared-intercept-breaking-fit}
and \eqref{eq:separate-intercept-breaking-fit} are both part of the analysis.

\subsection{Uncertainty Budget}
\noindent\textbf{Statistical errors.}
Statistical errors are obtained from blocked bootstrap samples. The block size
is chosen large enough to reduce residual autocorrelation effects in the saved
Markov-chain history. Derived observables such as Creutz ratios, Sommer scales,
and continuum-fit parameters are recomputed inside the bootstrap whenever
possible, so nonlinear error propagation is captured by the resampling.

\medskip
\noindent\textbf{Fit-window and interpolation systematics.}
Wilson-loop observables depend on analysis choices: the temporal range used for
effective-potential plateaus, the loop sizes included in Creutz-ratio summaries,
and the interpolation used to compare fixed physical distances. These choices
are varied to estimate systematic effects. For Creutz ratios, short distances
are most vulnerable to cutoff and breaking artifacts, while long distances are
most vulnerable to statistical noise.

\medskip
\noindent\textbf{Finite-volume and continuum-fit systematics.}
Finite-volume effects are monitored by comparing ensembles with similar
$\hat a$ and $\eps$ but different physical volumes when available. Continuum-fit
systematics are estimated by changing the fit ansatz, removing the coarsest
lattice spacing, and checking whether the extrapolation remains monotonic over
the fitted range. Non-monotonic behavior or strong shifts of the intercept under
reasonable fit variations are treated as evidence that the available lattice
spacings are not yet in a controlled scaling regime.

\section{Computational Environment}
The simulations were carried out on the QBig cluster. Most production runs were
performed on nodes 13--17 equipped with NVIDIA P100 GPUs. Additional and faster
runs were performed on nodes 18 and 19 equipped with NVIDIA A100 GPUs. The GPU
type affects the wall-clock runtime but not the generated Markov-chain ensemble,
provided that the same action, random seed handling, and update algorithm are
used.

\chapter{Results}
\label{chap:results}
\section{Thermalization and Data Quality}
Insert representative thermalization histories and autocorrelation estimates.

\section{Gauge-Invariant Baseline}
Present the $\eps=0$ results for $\wilsonloop$, $V(R)$, $\rzero$, and $\stringtension$. This establishes the reference continuum behavior.

\section{Dependence on Gauge-Breaking Strength}
\subsection{Observable Shifts at Fixed Coupling}
Report how each observable changes with $\eps_1$ at fixed $\beta$ and fixed
lattice geometry.

\subsection{Scaling of Shifts with Breaking Strength}
Fit or summarize whether the shifts are linear, quadratic, or nonlinear in the
breaking strength within the accessible range.

\subsection{Thresholds or Nonlinear Behavior}
Identify any breaking strengths for which the data no longer follow the
small-deformation pattern assumed in the scaling analysis.

\section{Continuum Limit at Fixed Breaking}
\subsection{Continuum Extrapolations for Each Breaking Strength}
Present continuum fits separately for each investigated breaking strength.

\subsection{Comparison with the Gauge-Invariant Limit}
Compare the extrapolated values with the $\eps=0$ baseline and state whether
the intercepts are statistically compatible.

\subsection{Stability Under Fit Variations}
Show how the continuum conclusions change when fit ranges, lattice spacings,
and systematic variants are modified.

\section{Combined Scaling Analysis}
Fit the dependence on both lattice spacing and gauge-breaking parameter. State whether the data support restoration of gauge invariance in the continuum limit.

\section{Finite-Volume and Systematic Checks}
Collect checks that are not part of the primary result but constrain the interpretation.

\chapter{Discussion}
\label{chap:discussion}
\section{Interpretation of the Continuum Limit}
Discuss whether small gauge breaking behaves like an irrelevant lattice artifact, a relevant deformation, or a source of uncontrolled continuum bias.

\section{Comparison with Theoretical Expectations}
Relate the numerical findings to the hypotheses from
Chapter~\ref{chap:small-breaking}.

\section{Limitations}
Discuss limits from statistics, accessible lattice spacings, finite volume, fit-systematic control, and the specific form of gauge breaking studied.

\section{Implications}
Explain what the result means for simulations or formulations where gauge invariance is only approximate.

\chapter{Conclusion and Outlook}
\label{chap:conclusion}
\section{Summary of Findings}
State the central quantitative conclusions.

\section{Answer to the Research Question}
Give the direct answer to whether the continuum limit is robust under the investigated small breaking of gauge invariance.

\section{Future Work}
Possible extensions:
\begin{itemize}
  \item Study additional forms of gauge-breaking terms.
  \item Increase statistics and extend to finer lattice spacings.
  \item Add gauge-variant diagnostics to directly measure symmetry restoration.
  \item Compare different gauge groups or dimensions.
\end{itemize}

\appendix

\chapter{Simulation Tables}
Provide complete tables of ensembles, including $\beta$, lattice size, breaking parameters, number of sweeps, thermalization cuts, and measured scales.

\chapter{Additional Plots}
Include supplementary Wilson-loop plateaus, bootstrap block-size scans, potential fits, and continuum-fit variants.

\chapter{Analysis Code and Reproducibility}
Document the exact command-line workflow, cache version, software environment, and data-selection rules used for the final analysis.

\chapter{Derivations}
Collect longer derivations that would interrupt the main text, such as Creutz-ratio identities or continuum-scaling arguments.

\bibliographystyle{unsrt}
\bibliography{references}

\end{document}
